---
title: "p8105_hw3_jy3270"
author: "Jingya Yu"
date: "2022-10-15"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set up library
- library we use
```{r}
library(tidyverse)
library(ggridges)
library(patchwork)
library(ggplot2)
library(p8105.datasets)
```
## Probelm 1
```{r}
data("instacart")

instacart = 
  instacart %>% 
  as_tibble(instacart)
```

#### Answer questions about the data

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row resprenting a single product from an instacart order. Variables include identifiers for user, order, and product; the order in which each product was added to the cart. There are several order-level variables, describing the day and time of the order, and number of days since prior order. Then there are several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. In total, there are `r instacart %>% select(product_id) %>% distinct %>% count` products found in `r instacart %>% select(user_id, order_id) %>% distinct %>% count` orders from `r instacart %>% select(user_id) %>% distinct %>% count` distinct users.

Below is a table summarizing the number of items ordered from aisle. In total, there are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered by far.

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

Next is a plot that shows the number of items ordered in each aisle. Here, aisles are ordered by ascending number of items.

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

Our next table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, and includes the number of times each item is ordered in your table.

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(desc(n)) %>%
  knitr::kable()
```

Finally is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. This table has been formatted in an untidy manner for human readers. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.

```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour) %>%
  knitr::kable(digits = 2)
```

## Problem 2
```{r}
# loading the data
accel_data <- read_csv(file = "./data/accel_data.csv") %>% janitor::clean_names()
```
- Load and tidy the data
```{r}
accel_tidy <- accel_data %>%  
  pivot_longer(cols = "activity_1":"activity_1440", 
               names_to = "minute", 
               values_to = "activity_count") %>%
  mutate(
    week_type = ifelse((day == "Saturday" | day == "Sunday"),
     "weekend", "weekday"),
  ) %>%
  select(week,day_id,day,week_type,everything())

head(accel_tidy)
```

This dataset contains `r ncol(accel_tidy)` variables, included week, day_id, day,
week_type, minute, and activity counts in each one-minute interval measured by 
of a 24-hour day starting from midnight, and collected `r nrow(accel_tidy)` 
observations in total.  

- create a total activity variable for each day  
```{r}
total_activity =
  accel_tidy %>%
  group_by(week, day) %>%
  summarise(total_activity_counts=sum(activity_count))

total_activity
```
I can't see any clear trend of activity count.

```{r}
total_activity %>%
  ggplot(aes(x = week, y = total_activity_counts, color = day, group = day)) +
  geom_point() +
  geom_line() + 
  labs(y = "total activity count by days in week", 
       title = "total activity count vs. week by days in week")
```

Still, there are no apparent trend when we focus on the total activity over the day

- single-panel plot that shows the 24-hour activity time course for each day
```{r}
accel_tidy %>%
  mutate(minute = as.integer(sub("activity_","",minute)),
         day = factor(day, levels =
                        c("Monday","Tuesday","Wednesday","Thursday","Friday",
                          "Saturday","Sunday"))) %>%
 ggplot(aes(x = minute, y = activity_count, color = day)) + 
    geom_point(alpha = 0.5) + 
    labs(
    title = "Activity time courses by day",
    x = "Minute",
    y = "Activity Count") +
  scale_x_continuous(
    breaks=c(60,120,180,240,300,360,420,480,540,600,660,720,780,840,900,960,1020,
             1080,1140,1200,1260,1320,1380,1440),
    labels=c("1:00 AM","2:00 AM","3:00 AM","4:00 AM","5:00 AM","6:00 AM",
             "7:00 AM","8:00 AM","9:00 AM","10:00 AM","11:00 AM","12:00 AM",
             "1:00 PM","2:00 PM","3:00 PM","4:00 PM","5:00 PM","6:00 PM","7:00 PM",
             "8:00 PM","9:00 PM","10:00 PM","11:00 PM","00:00 AM")
  ) +
  theme(axis.text.x=element_text(angle = 90))
```

## Problem 3
- load in data
```{r}
library(p8105.datasets)
noaa_df = data("ny_noaa")
```

This dataset contains `r nrow(noaa_df)` rows and `r ncol(noaa_df)` columns. Each row 
representing a single observation from a New York state weather station from
January 1, 1981 to December 31, 2010. Variables include weather station ID (id), 
date of observation (date), precipitation in tenths of mm (prcp), snowfall in mm (snow), 
snow depth in mm (snwd), maximum temperature in tenths of degrees C (tmax), 
and minimum temperature in tenths of degrees C (tmin). Since each weather station 
may collect only a subset of these variables, the resulting dataset contains 
substantial missing data.
